# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Data table is used to interact with the data to make operating over particular dates and intervals easier.

The activity data is read in and a total steps per day is calculated. Additionally, average steps per interval across days, is derived. NAs are treated as zero for this purpose.

Additionally summary metrics are computed for the steps per day and steps per interval.

```{r,results='hide'}
library("data.table")
library("ggplot2")
activityData <- data.table(read.csv("activity.csv"))

activityData[,stepsDay := sum(steps), by=date]
activityData[,stepsInterval :=mean(steps, na.rm=TRUE), by=interval]

meanSteps = mean(activityData$stepsDay, na.rm=TRUE)
medianSteps = median(activityData$stepsDay, na.rm=TRUE)

maxSteps = max(activityData$stepsInterval)

```


## What is mean total number of steps taken per day?

A histogram is generated of the steps taken per day, illustrating the spread of the data.

The mean number of steps per day is `r meanSteps` and the median steps per day is `r medianSteps`.

````{r}
qplot(stepsDay, data=activityData, binwidth=1000, xlab="steps per day")

```


## What is the average daily activity pattern?

Looking at the different 5-minute collection intervals throughout the day, we see the following average pattern of activity for a day.

```{r}

```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
